\documentclass[a4paper,14pt, oneside]{extarticle} 
\usepackage{cmap}                     % Поддержка поиска русских слов в PDF (pdflatex), ЗАГРУЖАТЬ САМЫМ ПЕРВЫМ!
\usepackage[unicode=true]{hyperref} % гиперрсылки. в том числе имим становятся обычные ссылки и оглавления.
\usepackage[a4paper,height=25cm,hmargin={3cm,1.5cm},vmargin={2.5cm,2.5cm}]{geometry} % меняет размеры прямо в процессе
%\usepackage{geometry}
%\geometry{left=3cm}% левое поле
%\geometry{right=1.5cm}% правое поле
%\geometry{top=2.5cm}% верхнее поле
%\geometry{bottom=2.5cm}% нижнее поле
\usepackage{mathtext}                 % Поддержка кириллицы в математическом режиме. ЗАГРУЖАТЬ ДО fontenc
%\usepackage[warn]{mathtext}          % замена предыдущему, только выводятся предупреждения об кириллице в формулах     
\usepackage[T2A]{fontenc}          % для использования русской раскладки необходимо
\usepackage[koi8-r,cp1251]{inputenc}         % набить данные три строчки
\usepackage[english]{babel}   % может быть использована и другая кодировка cp866 вместо cp1251
%\renewcommand{\theenumi}{\asbuk{enumi}} % чтобы enumerate русские буквы в перечнях использовал на верхнем уровне
\usepackage{amsmath,amssymb, amsxtra}
\usepackage{nccmath}                  % расширяет пакет amsmath
\usepackage{epsfig}                   % все остальные пакеты использую для корректной вставки графики
\usepackage{graphicx}                 % и т.п.
%\usepackage{graphics}
\usepackage{nccfloats}
\usepackage{color}
\usepackage{cite}    % позволяет выводить списки не [1, 2, 3], а [1-10]
%\usepackage{caption2}                % чтобы в подрисуночной подписи
%\renewcommand{\captionlabeldelim}{.} % ... после слова рис. ставилась точка
%\usepackage[font=small,labelfont=bf]{caption} 

% Редактируем стиль заголовков таблиц и рисунков
% nooneline --- чтобы сработало raggedright
\usepackage[format=plain, labelsep=period, labelfont=bf, textfont=small]{caption} 
%\captionsetup[table]{name=Табл.}
%\captionsetup[figure]{justification=centering}

%\usepackage{ccaption} % другой способ, чтобы в подрисуночной подписи
%\captiondelim{. }     % ...  после слова рис. ставилась точка
%\captionnamefont{\bf}
%\captiontitlefont{\small}
%Включаем символы Upgreek
\usepackage{upgreek}
\usepackage{afterpage}
\usepackage{placeins}
\usepackage{indentfirst} % Красная строка в первом абзаце 
%\usepackage[misccorr]
\usepackage[shortcuts,cyremdash]{extdash} % переносы в составных словах, дефисы и тире (из ncc-tools)
\usepackage[thinc,italic]{esdiff} % упрощает написание частных и прочих производных. Без italic d не будет курсивом
%\usepackage{commath} % почти тоже самое
\usepackage{shadow} % боксы красивые
\usepackage{xtab} % таблицы,с возможностью разбиения на страницы,с возможностью подстрчного примечания на каждой странице с таблицей
%\usepackage{enumerate} % расширение возможностей перечней
\usepackage{lscape}    % меняем ориентацию текста на 90 град...
\usepackage{pdflscape} % ... для этого необходимый текс помещаем в окружение landscape (например, таблтицу)
\usepackage{ncccomma} % используя этот пакет, запятая в 3,14 будет считать разделителем разрядов, а не десятков
\usepackage{mathrsfs} % чтобы писать всякие гамильтонианы
%\usepackage{dsfont} % \mathds
\usepackage{yfonts} % готика в математическом режиме
\usepackage[integrals]{wasysym} % чтобы интегралы были прямыми как в русском стиле
\usepackage{multirow} % ячейка на несколько строк
\usepackage[stable,perpage]{footmisc} % подстрочные примечания. с опцией perpage нумерация сносок при переходе на новую страницу опять начинается с единицы
% stable -- http://www.tex.ac.uk/cgi-bin/texfaq2html?label=ftnsect -- чтобы можно было использовать сноски в заголовках
\usepackage{prettyref} % пакет позволяет внешний вид ссылки в зависимости от префикса. к примеру, \newrefformat{eq}{формула \No$\sim$\ref{#1}}. Тогда на выходе получим \prettyref{eq:1} -> формула № 1.
\usepackage{titlesec}


\titleformat{\section}[block]{\Large\bfseries\filcenter}{}{1em}{}




\begin{document}
\newpage
\clearpage

{\raggedleft\textbf{Solodovnikov Mikhail,\\MiFIT 2013}

}

\section*{Measurment}

\subsection*{Defenition}

\textbf{Measurement} (from Old French, \textit{mesurement}) is the assignment of numbers to objects or events. It is a cornerstone of most natural sciences, technology, economics, and quantitative research in other social sciences.

Any measurement can be judged by the following meta--measurement criteria values: level of measurement (which includes magnitude), dimensions (units), and uncertainty. They enable comparisons to be done between different measurements and reduce confusion. Even in cases of clear qualitative similarity or difference, increased precision through quantitative measurement is often preferred in order to aid in replication. For example, different colours may be operationalized based either on wavelengths of light or (qualitative) terms such as "green" and "blue" which are often interpreted differently by different people. The science of measurement is called metrology.%\footnote{\href{http://en.wikipedia.org/wiki/Measurement}{Wikipedia material}}.

\textbf{Metrology} is the science of measurement. Metrology includes all theoretical and practical aspects of measurement. The word comes from Greek $\mu\varepsilon\tau\rho o\nu$ (metron), "measure" + "$\lambda o\gamma o\varsigma$" (logos), amongst others meaning "speech, oration, discourse, quote, study, calculation, reason". In Ancient Greek the term $\mu\varepsilon\tau\rho\lambda o\gamma i\alpha$ (metrologia) meant "theory of ratios".%\footnote{\href{http://en.wikipedia.org/wiki/Metrology}{Wikipedia material}}.

\subsection*{History of Measurment}

Weights and measures have taken a great variety of forms over the course of history, from simple informal expectations in barter transactions to elaborate state and supranational systems that integrate measures of many different kinds. Weights and measures from the oldest societies can often be inferred at least in part from archeological specimens, often preserved in museums. The comparison of the dimensions of buildings with the descriptions of contemporary writers is another source of information. An interesting example of this is the comparison of the dimensions of the Greek Parthenon with the description given by Plutarch from which a fairly accurate idea of the size of the Attic foot is obtained. Because of the comparative volume of artifacts and documentation, we know much more about the state--sanctioned measures of large, advanced societies than we do about those of smaller socieities or about the informal measures that often coexisted with official ones throughout history. In some cases, we have only plausible theories and we must sometimes select the interpretation to be given to the evidence.

\subsubsection*{Earliest Known Systems}

By studying the evidence given by all available sources, and by correlating the relevant facts, we obtain some idea of the origin and development of the units. We find that they have changed more or less gradually with the passing of time in a complex manner because of a great variety of modifying influences. It is possible to group official measurement systems for large societies into historical systems that are relatively stable over time, including: the Babylonian system, the Egyptian system, the Phileterian system of the Ptolemaic age, the Olympic system of Greece, the Roman system, the British system, and the metric system.

The inhabitants of the Indus Valley Civilization (3000 --- 1500 BC. Mature period 2600 --- 1900 BC) developed a sophisticated procedure of standardization, using weights and measures, evident by the excavations made at the Indus valley sites. This technical standardization enabled gauging devices to be effectively used in angular measurement and measurement for construction. Calibration was also found in measuring devices along with multiple subdivisions in case of some devices.

The earliest known uniform systems of weights and measures seem all to have been created at some time in the 4th and 3rd millennia BC among the ancient peoples of Egypt, Mesopotamia and the Indus Valley, and perhaps also Elam (in Iran) as well. The most astounding of these ancient systems was perhaps that of the Indus Valley Civilization (ca. 2600 BC). The Indus Valley peoples achieved great accuracy in measuring length, mass, and time. Their measurements were extremely precise since their smallest division, which is marked on an ivory scale found in Lothal, was approximately 1.704 mm (1/16 inch), the smallest division ever recorded on a scale of the Bronze Age. The decimal system was thought to be used but in fact, feet and inches are a more accurate indication of the measure used. The measuring devices found in the Indus Valley actually meet the standards used in South Indian architecture and engineering which are feet and inches. In Tamil Nadu, the inch is used in architectural forms per traditional Vaastu Shastras. Traditional architectural forms are built to 1/16 of an inch which is the smallest unit of measure found in the Indus Valley. In addition, the "yard sticks" used in Indus Valley (and in Tamil Nadu) measure exactly 33 inches and are divided into 24 units of 1$\frac{3}{8}$ inches. The 1$\frac{3}{8}$ inch is considered to be an angula. A unit of measure 33 inches long is called a hasta, kishku or muzam. A half hasta was also found in the Indus Valley measuring exactly 16$\frac{1}{2}$ inches (12 angula at 1$\frac{3}{8}$ inches). This measure is currently used by traditional architects such as Dr. V. Ganapati Sthapati who comes from an ancient tradition of Shilpis. The discovery of the relationship between Indus Valley measurement (using feet and inches rather than metric measure) and current and traditional Indian architecture per Vaastu Shastras was made by Dr. Jessie J. Mercay, Chancellor of The American University of Mayonic Science and Technology.

Weights were based on units of 0.05, 0.1, 0.2, 0.5, 1, 2, 5, 10, 20, 50, 100, 200, and 500, with each unit weighing approximately 28 grams, similar to the English ounce or Roman uncia, and smaller objects were weighed in similar ratios with the units of 0.871.

Other systems were based on the use of parts of the body and the natural surroundings as measuring instruments. Ancient Indian measurements related to the body are corelated to the finger measure of 1$\frac{3}{8}$ inch. This measure is found throughout the human body in increments. It is the measure used to build ancient temples and is precisely related to the Indus Valley measuring devices. Early Babylonian and Egyptian records and the Bible indicate that length was first measured with the forearm, hand, or finger and that time was measured by the periods of the sun, moon, and other heavenly bodies. When it was necessary to compare the capacities of containers such as gourds or clay or metal vessels, they were filled with plant seeds which were then counted to measure the volumes. When means for weighing were invented, seeds and stones served as standards. For instance, the carat, still used as a unit for gems, was derived from the carob seed.

\subsubsection*{History of Units}

\paragraph*{Units of Length}

The Egyptian cubit, the Indus Valley units of length referred to above and the Mesopotamian cubit were used in the 3rd millennium BC and are the earliest known units used by ancient peoples to measure length. The units of length used in ancient India included the dhanus (bow), the krosa (cry, or cow--call) and the yojana (stage).

The common cubit was the length of the forearm from the elbow to the tip of the middle finger. It was divided into the span of the hand (one--half cubit), the palm or width of the hand (one sixth), and the digit or width of the middle finger (one twenty--fourth) and the span or the length between the tip of little finger to the tip of the thumb. The Sacred Cubit, which was a standard cubit enhanced by an extra span --- thus 7 spans or 28 digits long --- was used in constructing buildings and monuments and in surveying in ancient Egypt. The inch, foot, and yard evolved from these units through a complicated transformation not yet fully understood. Some believe they evolved from cubic measures; others believe they were simple proportions or multiples of the cubit. In whichever case, the Greeks and Romans inherited the foot from the Egyptians. The Roman foot ($\sim$296 mm) was divided into both 12 unciae (inches) ($\sim$24.7 mm) and 16 digits ($\sim$18.5 mm). The Romans also introduced the mille passus (1000 paces) or double steps, the pace being equal to five Roman feet ($\sim$1480 mm). The Roman mile of 5000 feet (1480 m) was introduced into England during the occupation. Queen Elizabeth I (reigned from 1558 to 1603) changed, by statute, the mile to 5280 feet ($\sim$1609 m) or 8 furlongs, a furlong being 40 rod (unit)s ($\sim$201 m) of 5.5 yards ($\sim$5.03 m) each.

The introduction of the yard (0.9144 m) as a unit of length came later, but its origin is not definitely known. Some believe the origin was the double cubit, others believe that it originated from cubic measure. Whatever its origin, the early yard was divided by the binary method into 2, 4, 8, and 16 parts called the half-yard, span, finger, and nail. The association of the yard with the "gird" or circumference of a person's waist or with the distance from the tip of the nose to the end of the thumb of King Henry I (reigned 1100–1135) are probably standardizing actions, since several yards were in use in Britain. There were also Rods, Poles and Perches for measurements of length.

\paragraph*{Typographical Units}

The use in typography of the point, which was one twelfth of a ligne in the customary French measures of the Ancien Regime, as a unit for measuring print type predates the metric system. Various typographic points were defined, including the Truchet point by Sebastien Truchet (1657 --- 1729), the Fournier point by Pierre Simon Fournier in 1737 and the Didot point by the Didot brothers, Francois Ambroise and Pierre Francois, in 1755, or 1767, or 1783, which was exactly two Truchet points. The point was first used in the United States in 1878 or 1872 by a Chicago type foundry (Marder, Luse, \& Co.).

The point was defined as 1/12 of a pica consisting of 1/6 of an inch truncated to thousandths of an inch (1/12 of 0.166 inches) in the USA in 1886 and was quickly adopted by Britain and its colonies. This number has been approximated in various ways depending on the technology, prejudices, and information available to engineers involved in the production of typesetting machinery since then. Common approximations in use are 1/72.27 inch (now equal to 0.3514598 mm) and simply 1/72 of an inch. Other typographic points are used in other countries.

The pica in any of several related systems measures 12 points.

\paragraph*{Units of mass}

The grain was the earliest unit of mass and is the smallest unit in the apothecary, avoirdupois, Tower, and troy systems. The early unit was a grain of wheat or barleycorn used to weigh the precious metals silver and gold. Larger units preserved in stone standards were developed that were used as both units of mass and of monetary currency. The pound was derived from the mina used by ancient civilizations. A smaller unit was the shekel, and a larger unit was the talent. The magnitude of these units varied from place to place. The Babylonians and Sumerians had a system in which there were 60 shekels in a mina and 60 minas in a talent. The Roman talent consisted of 100 libra (pound) which were smaller in magnitude than the mina. The troy pound ($\sim$373.2 g) used in England and the United States for monetary purposes, like the Roman pound, was divided into 12 ounces, but the Roman uncia (ounce) was smaller. The carat is a unit for measuring gemstones that had its origin in the carob seed, which later was standardized at 1/144 ounce and then 0.2 gram.

Goods of commerce were originally traded by number or volume. When weighing of goods began, units of mass based on a volume of grain or water were developed. For example, the talent in some places was approximately equal to the mass of one cubic foot of water. Was this a coincidence or by design? The diverse magnitudes of units having the same name, which still appear today in our dry and liquid measures, could have arisen from the various commodities traded. The larger avoirdupois pound for goods of commerce might have been based on volume of water which has a higher bulk density than grain. For example, the Egyptian hon was a volume unit about 11 per cent larger than a cubic palm and corresponded to one mina of water. It was almost identical in volume to the present U.S. pint ($\sim$473 mL).

The stone, quarter, hundredweight, and ton were larger units of mass used in Britain. Today only the stone continues in customary use for measuring personal body weight. The present stone is 14 pounds ($\sim$6.35 kg), but an earlier unit appears to have been 16 pounds ($\sim$7.25 kg). The other units were multiples of 2, 8, and 160 times the stone, or 28, 112, and 2240 pounds ($\sim$12.7 kg, 50.8 kg, 1016 kg), respectively. The hundredweight was approximately equal to two talents. The ton of 2240 pounds is called the "long ton". The "short ton" is equal to 2000 pounds ($\sim$907 kg). A tonne (t) is equal to 1000 kg.

\paragraph*{Units of time and angle}

The division of the circle into 360 degrees and the day into hours, minutes, and seconds can be traced to the Babylonians who had sexagesimal system of numbers. The 360 degrees may have been related to a year of 360 days. Many other systems of measurement divided the day differently; other calendars divided the year differently.

\subsection*{Modern Metric Systems, Units, Tools}

\subsubsection*{Imperial and US Customary Systems}

Before SI units were widely adopted around the world, the British systems of English units and later imperial units were used in Britain, the Commonwealth and the United States. The system came to be known as U.S. customary units in the United States and is still in use there and in a few Caribbean countries. These various systems of measurement have at times been called foot--pound--second systems after the Imperial units for length, weight and time even though the tons, hundredweights, gallons, and nautical miles, for example, are different for the U.S. units. Many Imperial units remain in use in Britain, which has officially switched to the SI system --- with a few exceptions such as road signs, which are still in miles. Draught beer and cider must be sold by the imperial pint, and milk in returnable bottles can be sold by the imperial pint. Many people measure their height in feet and inches and their weight in stone and pounds, to give just a few examples. Imperial units are used in many other places, for example, in many Commonwealth countries that are considered metricated, land area is measured in acres and floor space in square feet, particularly for commercial transactions (rather than government statistics). Similarly, gasoline is sold by the gallon in many countries that are considered metricated.

\subsubsection*{Metric System}

The metric system is a decimal systems of measurement based on its units for length, the metre and for mass, the kilogram. It exists in several variations, with different choices of base units, though these do not affect its day--to--day use. Since the 1960s, the International System of Units (SI) is the internationally recognised metric system. Metric units of mass, length, and electricity are widely used around the world for both everyday and scientific purposes.

The metric system features a single base unit for many physical quantities. Other quantities are derived from the standard SI units. Multiples and fractions of the units are expressed as Powers of 10 of each unit. Unit conversions are always simple because they are in the ratio of ten, one hundred, one thousand, etc., so that convenient magnitudes for measurements are achieved by simply moving the decimal place: 1.234 metres is 1234 millimetres or 0.001234 kilometres. The use of fractions, such as 2/5 of a metre, is not prohibited, but uncommon. All lengths and distances, for example, are measured in metres, or thousandths of a metre (millimetres), or thousands of metres (kilometres). There is no profusion of different units with different conversion factors as in the Imperial system which uses, for example, inches, feet, yards, fathoms, rods.

\subsubsection*{International System of Units}

The International System of Units (abbreviated as SI from the French language name \textit{Systeme International d'Unites}) is the modern revision of the metric system. It is the world's most widely used system of units, both in everyday commerce and in science. The SI was developed in 1960 from the metre--kilogram--second (MKS) system, rather than the centimetre--gram--second (CGS) system, which, in turn, had many variants. During its development the SI also introduced several newly named units that were previously not a part of the metric system.

There are two types of SI units, base units and derived units. Base units are the simple measurements for time, length, mass, temperature, amount of substance, electric current and light intensity. Derived units are constructed from the base units, for example, the Watt, i.e. the unit for power, is defined from the base units as $m^2 \cdot kg\cdot s^{-3}$. Other physical properties may be measured in compound units, such as material density, measured in $kg/m^3$.

\paragraph*{Converting prefixes}
The SI allows easy multiplication when switching among units having the same base but different prefixes. To convert from metres to centimetres it is only necessary to multiply the number of metres by 100, since there are 100 centimetres in a metre. Inversely, to switch from centimetres to metres one multiplies the number of centimetres by 0.01 or divide centimetres by 100.

\subsubsection*{Length}

A ruler or rule is a tool used in, for example, geometry, technical drawing, engineering, and carpentry, to measure lengths or distances or to draw straight lines. Strictly speaking, the ruler is the instrument used to rule straight lines and the calibrated instrument used for determining length is called a measure, however common usage calls both instruments rulers and the special name straightedge is used for an unmarked rule. The use of the word measure, in the sense of a measuring instrument, only survives in the phrase tape measure, an instrument that can be used to measure but cannot be used to draw straight lines. As can be seen in the photographs on this page, a two--metre carpenter's rule can be folded down to a length of only 20 centimetres, to easily fit in a pocket, and a five--metre--long tape measure easily retracts to fit within a small housing.


\subsubsection*{Time}

The official SI definition of the second is as follows:\\
The second is the duration of 9,192,631,770 periods of the radiation corresponding to the transition between the two hyperfine levels of the ground state of the caesium 133 atom.

Time is an abstract measurement we have invented in order to keep track of elemental changes over a non spatial continuum. It is denoted by numbers and/or named periods such as hours, days, weeks, months and years. It is an apparently irreversible series of occurrences within this non spatial continuum. It is also used to denote an interval between two relative points on this continuum.

\paragraph*{Some special names}
We also use some special names for some multiples of some units.\\
10 years = 1 decade; 100 years = 1 century; 1000 years = 1 millennium


\subsubsection*{Mass}

Mass refers to the intrinsic property of all material objects to resist changes in their momentum. Weight, on the other hand, refers to the downward force produced when a mass is in a gravitational field. In free fall, (no net gravitational forces) objects lack weight but retain their mass. The Imperial units of mass include the ounce, pound, and ton. The metric units gram and kilogram are units of mass.

One device for measuring weight or mass is called a weighing scale or, often, simply a scale. A spring scale measures force but not mass, a balance compares weight, both require a gravitational field to operate. Some of the most accurate instruments for measuring weight or mass are based on load cells with a digital read--out, but require a gravitational field to function and would not work in free fall.

\paragraph*{Some special names}100 kilograms = 1 quintal; 1000 kilogram = 1 metric tonne;

\subsection*{Uncertainty Principle}

In quantum mechanics, the uncertainty principle is any of a variety of mathematical inequalities asserting a fundamental limit to the precision with which certain pairs of physical properties of a particle known as complementary variables, such as position x and momentum p, can be known simultaneously. For instance, the more precisely the position of some particle is determined, the less precisely its momentum can be known, and vice versa. The formal inequality relating the standard deviation of position $\sigma_x$ and the standard deviation of momentum $\sigma_p$ was derived by Earle Hesse Kennard later that year and by Hermann Weyl in 1928,
\begin{equation*}
\sigma_{x}\sigma_{p} \geq \frac{\hbar}{2} ~,
\end{equation*} 
where $\hbar$ is the reduced Planck constant.

The original heuristic argument that such a limit should exist was given by Werner Heisenberg in 1927, after whom it is sometimes named the Heisenberg principle. This ascribes the uncertainty in the measurable quantities to the jolt--like disturbance triggered by the act of observation. Though widely repeated in textbooks, this physical argument is now known to be fundamentally misleading. While the act of measurement does lead to uncertainty, the loss of precision is less than that predicted by Heisenberg's argument; the formal mathematical result remains valid, however.

Historically, the uncertainty principle has been confused with a somewhat similar effect in physics, called the observer effect, which notes that measurements of certain systems cannot be made without affecting the systems. Heisenberg offered such an observer effect at the quantum level (see below) as a physical "explanation" of quantum uncertainty. It has since become clear, however, that the uncertainty principle is inherent in the properties of all wave--like systems, and that it arises in quantum mechanics simply due to the matter wave nature of all quantum objects. Thus, the uncertainty principle actually states a fundamental property of quantum systems, and is not a statement about the observational success of current technology. It must be emphasized that measurement does not mean only a process in which a physicist-observer takes part, but rather any interaction between classical and quantum objects regardless of any observer.

Since the uncertainty principle is such a basic result in quantum mechanics, typical experiments in quantum mechanics routinely observe aspects of it. Certain experiments, however, may deliberately test a particular form of the uncertainty principle as part of their main research program. These include, for example, tests of number--phase uncertainty relations in superconducting or quantum optics systems. Applications dependent on the uncertainty principle for their operation include extremely low noise technology such as that required in gravitational--wave interferometers.

\end{document}